---
title: "LC_default prediction"
author: Jialing
date: Jan 18, 2018
output: 
  html_document: 
    number_sections: true
    #toc:true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## data # 133889*151
```{r lc}
setwd("~/Desktop/MOOC/R/lending club")
loan = read.csv("LoanStats_securev1_2016Q1.csv", stringsAsFactors = FALSE, skip=1, header=TRUE)
loan0 = loan
```
## EDA
```{r eda}
head(loan)
str(loan)
summary(loan)
colnames(loan)
```

### missing Data: numeric(NA), character ("")
```{r eda1}
#summary.default(loan)
head(loan[,c("issue_d",'last_pymnt_d','next_pymnt_d')])
num.na=sort(sapply(loan, function(x) {sum(is.na(x))/dim(loan)[1]}),decreasing=TRUE)
num.na
num.NA1 = sort(sapply(loan, function(x) {length(which(x==""))}), decreasing=TRUE)
num.NA1
```
### Character null data not showing
```{r eda3}
head(loan[, c("issue_d","last_pymnt_d","next_pymnt_d")])
summary.default(loan)
length(which(loan$next_pymnt_d==""))
#loan$term
# please check the last two rows/observations (133888, 133889), these two obseravations are missing
```

```{r eda2}
library(VIM)
aggr_plot <- aggr(loan,col=c('white','red'), numbers=TRUE, sortVars=TRUE, labels=names(loan), cex.axis=.7,gap=3,ylab=c("Histogram of missing data","Pattern"))

#aggr_plot <- aggr(loan,col=c('navyblue','red'), labels=names(loan), cex.axis=.7,ylab=c("Histogram of missing data","Pattern"))
```
```{r lib}
# library zoo for time date 
library(zoo)
loan[,c("loan_status","next_pymnt_d")]
table(loan$loan_status)
loan$next_pymnt_d[1]
```

### Timestamp character
```{r eda30}
dim(subset(loan,next_pymnt_d==""))
#length(which(loan$next_pymnt_d==""))
with(subset(loan, next_pymnt_d==""),table(loan_status))
# Charged Off违约， 一直没有pay
```

delete loan_status=""
```{r eda4}
loan[which(loan$loan_status==""),]
loan=loan[-which(loan$loan_status==""),]
#subset(loan,loan_status=="")
#loan=subset(loan,loan_status!="")
```
###not considering those observations that 'next_pymnt_d'&'last_pymnt_d' are null
these loans will directly lead to "Charged Off or Default" status
```{r eda5}
with(subset(loan, last_pymnt_d==""),table(loan_status))
#charged off违约, loan很早以前就停止payment了, past due
#subset(loan, next_pymnt_d !=""&last_pymnt_d !="")
table(subset(loan, next_pymnt_d !=""&last_pymnt_d !="")$loan_status)
```
### Life Cycle
considering date difference (loan 'life cycle') instead of considering the date absolute value itself
```{r eda6}
head(loan$issue_d)
loan$issue_d_1=as.Date(as.yearmon(loan$issue_d,"%b-%Y"))
#loan$issue_d_1
loan$last_pymnt_d_1=as.Date(as.yearmon(loan$last_pymnt_d,"%b-%Y"))
loan$last_pymnt_from_issue=with(loan, last_pymnt_d_1-issue_d_1)
```
```{r eda7}
#loan$last_pymnt_from_issue
#plot(density(loan$last_pymnt_from_issue))
#typeof(loan$last_pymnt_from_issue)
table(subset(loan, last_pymnt_from_issue==0)$loan_status)
#table(loan$last_pymnt_from_issue)
```
Hard to know whether there is a linear relationship between last_pymnt_from_issue and loan_status, thus we will:
categorize/grouping/clustering "last_pymnt_from_issue":
0-3months: level one
4-6months: 
7-12months:
tradeoff: partition into more smaller groups (more levels, more features), or loss some information of the groups
```{r eda8}
#loan1$state_mean_int <-
#  ifelse(loan1$addr_state %in% names(int_state)[which(int_state <= quantile(int_state, 0.25))], #         'low', ifelse(loan1$addr_state %in% names(int_state)[which(int_state <= quantile(int_state, 0.5))],
#                       'lowmedium', ifelse(loan1$addr_state %in% names(int_state)[which(int_state <= quantile(int_state, 0.75))], 
#                                           'mediumhigh', 'high')))
#sort(table(loan1$state_mean_int))

loan$last_pymnt_from_issue_cat=with(loan,as.character(cut(as.numeric(last_pymnt_from_issue),c(-1,0,92,184,366,549,639,730))))
```
```{r eda9}
table(loan$last_pymnt_from_issue_cat)
sum(table(loan$last_pymnt_from_issue_cat))
sum(is.na(loan$last_pymnt_from_issue))
```
### manipulate cat missing value
```{r eda10}
loan$last_pymnt_from_issue[which(is.na(loan$last_pymnt_from_issue))]=0
loan$last_pymnt_from_issue_cat[which(is.na(loan$last_pymnt_from_issue_cat))]="no pymnt"
```
```{r eda11}
table(loan$last_pymnt_from_issue_cat)
```
```{r eda12}
by.pymnt.gap=with(loan, table(loan_status,last_pymnt_from_issue_cat))
by.pymnt.gap
```
```{r eda13}
by.pymnt.gap=by.pymnt.gap[c("Charged Off","Default","Late (31-120 days)", "Late (16-30 days)", "In Grace Period","Current","Fully Paid"),c("(-1,0]","(0,92]","(92,184]","(184,366]","(366,549]","(549,639]","(639,730]")]
by.pymnt.gap
```
piecewise linear

```{r eda14}
round(by.pymnt.gap*100/apply(by.pymnt.gap,1,sum),3)
```
## logistic model test
```{r ml}
loan$loan_status_binary=as.factor(ifelse(loan$loan_status %in% c("Current","Fully Paid"), "okay","past due"))
```
```{r ml1}
mod1=glm(loan_status_binary~last_pymnt_from_issue_cat,loan,family="binomial")
mod2=glm(loan_status_binary~last_pymnt_from_issue,loan,family="binomial")
summary(mod1)
summary(mod2)
```

# why the result is weird from mod1?
# this is complete or more accurately quasi complete separatable data.
# MLE for logistic regression doesn't exist in the case of complete or more accurately quasi.
# https://stats.stackexchange.com/questions/224863/understanding-complete-separation-for-logistic-regression/224864#224864

```{r ml2}
with(subset(loan,last_pymnt_from_issue_cat =='(-1,0]'),table(loan_status_binary))
```

# Let's start with feature processing.
```{r fe1}
#check Date columns, try string function, "issue_d", ""
str(loan)
date.cols=colnames(loan)[c(which(grepl("_d$",colnames(loan))),which(grepl("_date$",colnames(loan))))]
```
```{r fe2}
for(col_i in date.cols) {
  loan[,col_i]=as.Date(as.yearmon(loan[,col_i],'%b-%Y'))
}
```
```{r fe3}
str(loan)
```
```{r fe4}
date.cols
sort(table(loan$last_pymnt_d),decreasing = TRUE)
```

compare with issue_date, or current_date: random set, e.g. '2017-11-1'
```{r fe5}
loan$mths_since_last_credit_pull=as.integer(as.Date('2017-11-1')-loan$last_credit_pull_d)
summary(loan$mths_since_last_credit_pull)
```
```{r fe6}
head(subset(loan,mths_since_last_credit_pull==0)$last_credit_pull_d)
```
```{r fe7}
loan$mths_since_last_credit_pull[which(is.na(loan$mths_since_last_credit_pull))]=0
```

## Interview考function achieve cross validation
### create/define a function
```{r fe8}
TransformToLengthFromIssueDate = function(loan,col.name, new.col.name,other.level) {
  tmp=as.integer(loan[,col.name]-loan$issue_d)/30
  loan[,new.col.name]=ifelse(is.na(loan[,col.name]),other.level,
                          as.character(cut(tmp,c(min(tmp,na.rm=TRUE)-1,quantile(tmp,c(0.1,0.5,0.9),na.rm=TRUE),max(tmp,na.rm=TRUE)+1))))
  return(loan)
}
```
```{r fe9}
loan=TransformToLengthFromIssueDate(loan,'hardship_start_date','hardship_since_issue','no_hs')
loan=TransformToLengthFromIssueDate(loan,'settlement_date','settlement_since_issue','no_settle')
```

```{r fe10}
date.cols
zz=as.data.frame(date.cols)
loanq=loan[,-c(zz[[1]])]
#loan=loan[,-which(colnames(loan) %in% date.cols)]
```

```{r fe11}
num.value=sapply(loan,function(x){return(length(unique(x)))})
num.value
```

character
```{r fe12}

feature.w.many.levels=colnames(loan)[intersect(which(sapply(loan,function(x){return(is.character(x))})),which(num.value>=50))]
```

```{r fe13}
#head(loan$int_rate)
#loan$int_rate=as.numeric(sapply(strsplit(loan$int,"%"),'[',1))
#head(loan$int_rate)
#typeof(loan$int_rate)
```

"[" extrac element
```{r fe14}
loanq$int_rate=as.numeric(sapply(strsplit(loanq$int,"%"),'[',1))
loanq$revol_util <- as.numeric(sapply(strsplit(loanq$revol_util, '%'), '[', 1))
#typeof(as.numeric(sapply(strsplit(loan$revol_util,"%"),"[",1)))
```

### 1. select top 10 titles, see how much it covers, and "others""; 2. or use NLP extract information
```{r fe15}
#table(loan$emp_title)
sum(with(loan,emp_title==""))
sort(table(loan$emp_title),decreasing=TRUE)
```

```{r fe16}
str(loan)
```

missing value
```{r fe17}
num.NA=sort(sapply(loan, function(x){sum(is.na(x))}), decreasing=TRUE)
num.NA[1:20]
```

```{r fe18}
num.empty <- sapply(loan[, colnames(loan)[which(sapply(loan, function(x){return(is.character(x))}))]],
                    function(x){return(length(which(x == "")))})
num.empty[which(num.empty > 0)]
num.empty
```
```{r fe19}
str(loan0)
```
```{r fe20}
```
```{r fe21}
```
```{r fe22}
```
